---
title: "Week 5 Assignment"
author: "Ben Arancibia"
date: "September 30, 2014"
output: html_document
---

__*Question 1*__

Three questions that I want answer based on the data presented in the assignment are: \n

1: What are the total populations for each city so that the proportion for each response can be calculated.

2: What proportion of the respondents in each city voted yes or no?

3: What proportion of the respondents in each age group voted yes or no?

In order for questions 2 and 3 to be answered, you need the answer for question one.

__*Question 2*__
```{r}
require(tidyr)
require(plyr)
require(ggplot2)
```
Create a dataframe to save this data in its current messy state.

```{r}
response <- c('Yes', 'No')
young_edinburgh <- c(80100, 35900)
older_edinburgh <- c(143000, 214800)
young_glasgow <- c(99400, 150400)
older_glasgow <- c(43000, 207000)
```

Now create a tidy data set from the above messy state.
```{r}
cleaned_response <- data.frame(response, young_edinburgh,
                               older_edinburgh, young_glasgow,
                               older_glasgow)

head(cleaned_response)
```
